01 Calculating Travel Time for TTS16 OD

### 2023-11-08 data-raw/00-TTS16-and-Bound-Data-Preparation-OD.Rmd
Working on development branch to ensure reproducibility of data processing and adding new big od table

Load packages:
```{r load-packages}
library(readxl)
library(sf)
library(tidyverse)
library(units)
```

@@

Read the boundaries of the traffic analysis zones:
```{r}
ggh_taz <- st_read("TTS16-data-inputs/OD-by-FT-Employment/tts06_83_region.shp")
```

Project object:
```{r}
ggh_taz <- ggh_taz %>%
  st_transform(crs = 32617)
```

Prepare traffic analysis zones:
```{r}
ggh_taz <- ggh_taz %>%
  transmute(GTA06 = as.character(GTA06),
            AREA = st_area(geometry) %>% 
              set_units(km^2) %>% 
              drop_units())
```

@@

Read the boundaries of the traffic analysis zones:
```{r}
ggh_pd <- st_read("TTS16-data-inputs/Boundaries/tts06_pd_83.shp")
```

Project object:
```{r}
ggh_pd <- ggh_pd %>%
  st_transform(crs = 32617)
```

Adding region names to the object as they are not included in the origin data:
(http://dmg.utoronto.ca/pdf/tts/2016/2016TTS_Conduct.pdf):
```{r}
ggh_pd <- ggh_pd %>%
  mutate(REGION_name = 
           ifelse(REGION == 1, "Toronto", 
                  ifelse(REGION == 2, "Durham", 
                         ifelse(REGION == 3, "York", 
                                ifelse(REGION == 4, "Peel", 
                                       ifelse(REGION == 5, "Halton", 
                                              ifelse(REGION == 6, "Hamilton", 
                                                     ifelse(REGION == 11, "Niagara", 
                                                            ifelse(REGION == 12, "Waterloo", 
                                                                   ifelse(REGION == 13, "Guelph", 
                                                                          ifelse(REGION == 14, "Wellington", 
                                                                                 ifelse(REGION == 15, "Orangeville", 
                                                                                        ifelse(REGION == 16, "Barrie", 
                                                                                               ifelse(REGION == 17, "Simcoe", 
                                                                                                      ifelse(REGION == 18, "Kawartha Lakes", 
                                                                                                             ifelse(REGION == 19, "Peterborough City", 
                                                                                                                    ifelse(REGION == 20, "Peterborough County", 
                                                                                                                           ifelse(REGION == 21, "Orillia", 
                                                                                                                                  ifelse(REGION == 22, "Dufferin", 
                                                                                                                                         ifelse(REGION == 23, "Brantford", 
                                                                                                                                                ifelse(REGION == 24, "Brant", "NOOOO")))))))))))))))))))))
```

## Origin destination matrices for work and shop

Read trip tables. The query was for "trips" by zone of residence, zone of work, and trip purpose (Work):
```{r}
od_ft <- read_delim(file = "TTS16-data-inputs/OD-by-FT-Employment/tts-2016-OD-FT-Employment.txt", 
                      delim = "\t", 
                      col_names = FALSE)
```

@@

Slice the table to obtain the work trips (remove heading information):
```{r}
od_ft <- od_ft %>% 
  slice((idx+2):n())
```

Separate the zone identifiers and the trips, and convert to numeric:
```{r}
od_ft <- od_ft %>%
  separate(X1, into = c("Zones", "Persons"), sep = " (?=[^ ]+$)") %>%
  mutate(Zones = str_trim(Zones, side = "both")) %>%
  separate(Zones, into = c("Origin", "Destination"), sep = " (?=[^ ]+$)") %>%
  mutate(Origin = str_trim(Origin),
         Destination = str_trim(Destination),
         Persons = as.numeric(Persons))
```

@@

@@

The od table can be used to calculate the workers and jobs by traffic analysis zone:
```{r}
workers_ft <- od_ft %>%
  group_by(Origin) %>%
  summarize(workers = sum(Persons),
            .groups = "drop") %>%
  transmute(GTA06 = Origin,
            workers)

jobs_ft <- od_ft %>%
  group_by(Destination) %>%
  summarize(jobs = sum(Persons),
            .groups = "drop") %>%
  transmute(GTA06 = Destination,
            jobs)

Join workers and jobs to simple features object:
```{r}
ggh_taz <- ggh_taz %>%
  # Join workers
  left_join(workers_ft,
            by = "GTA06") %>%
  # Join jobs
  left_join(jobs_ft,
            by = "GTA06") %>%
  # Replace NAs with zeros
  mutate(workers = replace_na(workers, 0),
         jobs = replace_na(jobs, 0))
```

@@

@@

Attach the centroid point geometry to the origin object (workers) and the destination object (jobs)
```{r}
#workers (origins)
workers_ft_cent <- workers_ft %>% merge(ggh_taz_cent %>% dplyr::select(GTA06), by="GTA06") %>% st_sf()

work_origins <- cbind(workers_ft_cent, st_coordinates(st_transform(workers_ft_cent, crs = 4326))) %>%
  dplyr::rename(lon = "X", lat = "Y", id = "GTA06") %>% as.data.frame()  %>%
  dplyr::select(id, lon, lat) %>% as.data.frame()

@@

```{r}
#jobs (destinations)
jobs_ft_cent <- jobs_ft %>% merge(ggh_taz_cent %>% dplyr::select(GTA06), by="GTA06") %>% st_sf

job_destinations <- cbind(jobs_ft_cent, st_coordinates(st_transform(jobs_ft_cent, crs = 4326))) %>%
  dplyr::rename(lon = "X", lat = "Y", id = "GTA06") %>% as.data.frame()  %>%
  dplyr::select(id, lon, lat)

>>>>

Load packages:
```{r load-packages}
library(dplyr) # A Grammar of Data Manipulation
library(fitdistrplus) # Help to Fit of a Parametric Distribution to Non-Censored or Censored Data
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
#library(logspline) # not installed on this machine
library(readr) # not installed on this machine
library(sf) # Simple Features for R
library(stringr)
library(tidyr)
#library(tidyverse) # not installed on this machine
#library(tmap) # Thematic Maps
library(units) # Measurement Units for R Vectors
```

@@

Read the boundaries of the traffic analysis zones (retrieved from the TTS 2016):
```{r}
ggh_taz <- st_read("TTS16-data-inputs/OD-by-FT-Employment/tts06_83_region.shp")
```

Project object:
```{r}
ggh_taz <- ggh_taz |>
  st_transform(crs = 32617)
```

Prepare traffic analysis zones:
```{r}
ggh_taz <- ggh_taz |>
  transmute(GTA06 = as.character(GTA06),
            AREA = st_area(geometry) |> 
              set_units(km^2) |> 
              drop_units())
```

@@

Read the boundaries of the traffic analysis zones (retrieved from the TTS 2016):
```{r}
ggh_pd <- st_read("TTS16-data-inputs/Boundaries/tts06_pd_83.shp")
```

Project object:
```{r}
ggh_pd <- ggh_pd |>
  st_transform(crs = 32617)
```

Adding region names to the object as they are not included in the origin data:
(http://dmg.utoronto.ca/pdf/tts/2016/2016TTS_Conduct.pdf):
```{r}
ggh_pd <- ggh_pd |>
  mutate(REGION_name = case_when(REGION == 1 ~ "Toronto",
                                  REGION == 2 ~ "Durham",
                                  REGION == 3 ~ "York",
                                  REGION == 4 ~ "Peel",
                                  REGION == 5 ~ "Halton",
                                  REGION == 6 ~ "Hamilton",
                                  REGION == 11 ~ "Niagara",
                                  REGION == 12 ~ "Waterloo",
                                  REGION == 13 ~ "Guelph",
                                  REGION == 14 ~ "Wellington",
                                  REGION == 15 ~ "Orangeville",
                                  REGION == 16 ~ "Barrie",
                                  REGION == 17 ~ "Simcoe",
                                  REGION == 18 ~ "Kawartha Lakes",
                                  REGION == 19 ~ "Peterborough City",
                                  REGION == 20 ~ "Peterborough County", 
                                  REGION == 21 ~ "Orillia", 
                                  REGION == 22 ~ "Dufferin",
                                  REGION == 23 ~ "Brantford",
                                  REGION == 24 ~ "Brant"))
```

add 'REGION' from the planning distract object to the `ggh_taz`. 
```{r}
#buffering to clean up the edges of these objects -> there wre issues using st_join. 
ggh_taz <- st_buffer(ggh_taz, 
                     dist = 0)

ggh_pd <- st_buffer(ggh_pd, 
                    dist = 0)

ggh_taz <- st_join(ggh_taz, 
                   ggh_pd |>
                     transmute(REGION),
                   left = TRUE, 
                   largest = TRUE)
```

## CMA and CA boundaries

If there's interest in tying the TAZ to the Canadian Census, the boundaries and IDs for census metropolitan areas (CMA) and census ares (CA) are provided in this package as well. These areas are determined by Statistics Canada and

Read the boundaries of the traffic analysis zones (retrieved from the TTS 2016):

<!-- FILE NOT IN FOLDER -->
```{r}
ggh_cma <- st_read("TTS16-data-inputs/Boundaries/lcma000b16a_e.shp")
```

Project object and retain specific columns:
```{r}
ggh_cma <- ggh_cma |>
  st_transform(crs = 32617) |>
  transmute(CMAUID,
            CMANAME,
            CMATYPE)
```

add 'CMAUID' from the cma object to the `ggh_taz`. 
```{r}
#doing a st_join doesn't provide a clean solution, so I find the centroids of each TAZ, st_join them to the CMA/CA, and then join the resulting CMAUID property to the ggh_taz object
test_pts <-st_centroid(ggh_taz)
test<-st_join(test_pts, ggh_cma |> transmute(CMAUID), left = TRUE, join=st_within)
ggh_taz <-merge(ggh_taz, test |> transmute(GTA06, CMAUID)|> st_drop_geometry(), by=c("GTA06"))
```

## Origin destination matrices for place of residence and typical place work (by person)
Read exported TTS 2016 tables. The query was a cross-tabulation of "persons" by zone of residence and place of employment. The data was filtered to only include full-time not-remote workers and was retrieved from the TTS 2016:
```{r}
od_ft <- read_delim(file = "TTS16-data-inputs/OD-by-FT-Employment/tts-2016-OD-FT-Employment.txt", 
                      delim = "\t", 
                      col_names = FALSE)
```

@@

Slice the table to obtain the work trips (remove heading information):
```{r}
od_ft <- od_ft |> 
  slice((idx+2):n())
```

Separate the zone identifiers and the trips, and convert to numeric:
```{r}
od_ft <- od_ft |>
  separate(X1, into = c("Zones", "Persons"), sep = " (?=[^ ]+$)") |>
  mutate(Zones = str_trim(Zones, side = "both")) |>
  separate(Zones, into = c("Origin", "Destination"), sep = " (?=[^ ]+$)") |>
  mutate(Origin = str_trim(Origin),
         Destination = str_trim(Destination),
         Persons = as.numeric(Persons))
```

@@

## Trips taken (by mode) by worker from place of residence to job at place of work

Read exported TTs 2016 tables. The query was a cross-tablulation of "trips" by zone of residence and place of employment. The data was filtered to only include full-time not-remote working trip purposes by primary mode and was retrieved from the TTS 2016:

<!-- FILE DOES NOT EXIST IN DIRECTORY -->
```{r}
od_ft_trips <- read_delim(file = "TTS16-data-inputs/OD-by-FT-Employment/tts-2016-OD-FT-Employment-by_W_R_trip-byprimmode.txt", 
                      delim = "\t", 
                      col_names = FALSE)
```

Find the position in the table where the information on work trips begins:
```{r}
id_transit <- which(od_ft_trips$X1 == "TABLE    : mode_prime (Transit excluding GO rail)")
id_cycle <- which(od_ft_trips$X1 == "TABLE    : mode_prime (Cycle)")
id_car_driver <- which(od_ft_trips$X1 == "TABLE    : mode_prime (Auto driver)")
id_GO <- which(od_ft_trips$X1 == "TABLE    : mode_prime (GO rail only)")
id_GO_transit <- which(od_ft_trips$X1 == "TABLE    : mode_prime (Joint GO rail and local transit)")
id_motorcycle <- which(od_ft_trips$X1 == "TABLE    : mode_prime (Motorcycle)")
id_other <- which(od_ft_trips$X1 == "TABLE    : mode_prime (Other)")
id_car_pass <- which(od_ft_trips$X1 == "TABLE    : mode_prime (Auto passenger)")
id_schoolbus <- which(od_ft_trips$X1 == "TABLE    : mode_prime (School bus)")
id_taxi <- which(od_ft_trips$X1 == "TABLE    : mode_prime (Taxi passenger)")
id_rideshare <- which(od_ft_trips$X1 == "TABLE    : mode_prime (Paid rideshare)")
id_walk <- which(od_ft_trips$X1 == "TABLE    : mode_prime (Walk)")
id_end <- od_ft_trips$X1 |> length()
```

Slice the table to obtain the work trips by primary mode (there are 12 possible modes, so 12 new tables) (remove heading information and other modes):
```{r}
od_ft_trips_transit <- od_ft_trips |> 
  slice((id_transit+3):id_cycle-1)

od_ft_trips_cycle <- od_ft_trips |> 
  slice((id_cycle+3):id_car_driver-1)

od_ft_trips_car_driver <- od_ft_trips |> 
  slice((id_car_driver+3):id_GO-1)

od_ft_trips_GO <- od_ft_trips |> 
  slice((id_GO+3):id_GO_transit-1)

od_ft_trips_GO_transit <- od_ft_trips |> 
  slice((id_GO_transit+3):id_motorcycle-1)

od_ft_trips_motorcycle <- od_ft_trips |> 
  slice((id_motorcycle+3):id_other-1)

od_ft_trips_other <- od_ft_trips |> 
  slice((id_other+3):id_car_pass-1)

od_ft_trips_car_pass <- od_ft_trips |> 
  slice((id_car_pass+3):id_schoolbus-1)

od_ft_trips_schoolbus <- od_ft_trips |> 
  slice((id_schoolbus+3):id_taxi-1)

od_ft_trips_taxi<- od_ft_trips |> 
  slice((id_taxi+3):id_rideshare-1)

od_ft_trips_rideshare <- od_ft_trips |> 
  slice((id_rideshare+3):id_walk-1)

od_ft_trips_walk <- od_ft_trips |> 
  slice((id_walk+2):id_end)
```

Let's consolidate the 12 tables into 1 table, based on origin - destination and clean up the table. If there's no trip using that mode, a 0 is added. The following are the codes:

- T_LT - Local transit excluding GO rail
- T_LTGO - Local transit and GO
- T_GO - GO rail excluding local transit
```{r}
od_ft_trips_transit <- od_ft_trips_transit |>
  separate(X1, into = c("Zones", "T_LT"), sep = " (?=[^ ]+$)") |>
  mutate(Zones = str_trim(Zones, side = "both")) |>
  separate(Zones, into = c("Origin", "Destination"), sep = " (?=[^ ]+$)") |>
  mutate(Origin = str_trim(Origin),
         Destination = str_trim(Destination),
         T_LT = as.numeric(T_LT))

od_ft_trips_GO_transit <- od_ft_trips_GO_transit |>
  separate(X1, into = c("Zones", "T_LTGO"), sep = " (?=[^ ]+$)") |>
  mutate(Zones = str_trim(Zones, side = "both")) |>
  separate(Zones, into = c("Origin", "Destination"), sep = " (?=[^ ]+$)") |>
  mutate(Origin = str_trim(Origin),
         Destination = str_trim(Destination),
         T_LTGO = as.numeric(T_LTGO))

od_ft_trips_GO <- od_ft_trips_GO |>
  separate(X1, into = c("Zones", "T_GO"), sep = " (?=[^ ]+$)") |>
  mutate(Zones = str_trim(Zones, side = "both")) |>
  separate(Zones, into = c("Origin", "Destination"), sep = " (?=[^ ]+$)") |>
  mutate(Origin = str_trim(Origin),
         Destination = str_trim(Destination),
         T_GO = as.numeric(T_GO))
```

- BC - Cycle
- W - Walk
```{r}
od_ft_trips_cycle <- od_ft_trips_cycle |>
  separate(X1, into = c("Zones", "BC"), sep = " (?=[^ ]+$)") |>
  mutate(Zones = str_trim(Zones, side = "both")) |>
  separate(Zones, into = c("Origin", "Destination"), sep = " (?=[^ ]+$)") |>
  mutate(Origin = str_trim(Origin),
         Destination = str_trim(Destination),
         BC = as.numeric(BC))

od_ft_trips_walk <- od_ft_trips_walk |>
  separate(X1, into = c("Zones", "W"), sep = " (?=[^ ]+$)") |>
  mutate(Zones = str_trim(Zones, side = "both")) |>
  separate(Zones, into = c("Origin", "Destination"), sep = " (?=[^ ]+$)") |>
  mutate(Origin = str_trim(Origin),
         Destination = str_trim(Destination),
         W = as.numeric(W))
```

- C_DPC - Driver private car 
- C_PPC - Passenger private car
- C_TA - Taxi
- C_RS - Rideshare
- C_SB - School bus
```{r}
od_ft_trips_car_driver <- od_ft_trips_car_driver |>
  separate(X1, into = c("Zones", "C_DPC"), sep = " (?=[^ ]+$)") |>
  mutate(Zones = str_trim(Zones, side = "both")) |>
  separate(Zones, into = c("Origin", "Destination"), sep = " (?=[^ ]+$)") |>
  mutate(Origin = str_trim(Origin),
         Destination = str_trim(Destination),
         C_DPC = as.numeric(C_DPC))

od_ft_trips_car_pass <- od_ft_trips_car_pass |>
  separate(X1, into = c("Zones", "C_PPC"), sep = " (?=[^ ]+$)") |>
  mutate(Zones = str_trim(Zones, side = "both")) |>
  separate(Zones, into = c("Origin", "Destination"), sep = " (?=[^ ]+$)") |>
  mutate(Origin = str_trim(Origin),
         Destination = str_trim(Destination),
         C_PPC = as.numeric(C_PPC))

od_ft_trips_taxi <- od_ft_trips_taxi |>
  separate(X1, into = c("Zones", "C_TA"), sep = " (?=[^ ]+$)") |>
  mutate(Zones = str_trim(Zones, side = "both")) |>
  separate(Zones, into = c("Origin", "Destination"), sep = " (?=[^ ]+$)") |>
  mutate(Origin = str_trim(Origin),
         Destination = str_trim(Destination),
         C_TA = as.numeric(C_TA))

od_ft_trips_rideshare <- od_ft_trips_rideshare |>
  separate(X1, into = c("Zones", "C_RS"), sep = " (?=[^ ]+$)") |>
  mutate(Zones = str_trim(Zones, side = "both")) |>
  separate(Zones, into = c("Origin", "Destination"), sep = " (?=[^ ]+$)") |>
  mutate(Origin = str_trim(Origin),
         Destination = str_trim(Destination),
         C_RS = as.numeric(C_RS))

od_ft_trips_schoolbus <- od_ft_trips_schoolbus |>
  separate(X1, into = c("Zones", "C_SB"), sep = " (?=[^ ]+$)") |>
  mutate(Zones = str_trim(Zones, side = "both")) |>
  separate(Zones, into = c("Origin", "Destination"), sep = " (?=[^ ]+$)") |>
  mutate(Origin = str_trim(Origin),
         Destination = str_trim(Destination),
         C_SB = as.numeric(C_SB))
```

- M - Motorcycle
- O - Other
```{r}
od_ft_trips_motorcycle<- od_ft_trips_motorcycle |>
  separate(X1, into = c("Zones", "M"), sep = " (?=[^ ]+$)") |>
  mutate(Zones = str_trim(Zones, side = "both")) |>
  separate(Zones, into = c("Origin", "Destination"), sep = " (?=[^ ]+$)") |>
  mutate(Origin = str_trim(Origin),
         Destination = str_trim(Destination),
         M = as.numeric(M))

od_ft_trips_other <- od_ft_trips_other |>
  separate(X1, into = c("Zones", "O"), sep = " (?=[^ ]+$)") |>
  mutate(Zones = str_trim(Zones, side = "both")) |>
  separate(Zones, into = c("Origin", "Destination"), sep = " (?=[^ ]+$)") |>
  mutate(Origin = str_trim(Origin),
         Destination = str_trim(Destination),
         O = as.numeric(O))
```

And now, consolidate all 12 tables into 1 table! with 12 columns for mode trips numbers + the origin and destination columns (those are GTA06 TAZ IDs):
```{r}
od_ft_trips <- full_join(od_ft_trips_transit, od_ft_trips_GO_transit, by=c("Origin", "Destination")) |>
  full_join(od_ft_trips_GO, by=c("Origin", "Destination")) |>
  full_join(od_ft_trips_cycle, by=c("Origin", "Destination")) |>
  full_join(od_ft_trips_walk, by=c("Origin", "Destination")) |>
  full_join(od_ft_trips_car_driver, by=c("Origin", "Destination")) |>
  full_join(od_ft_trips_car_pass, by=c("Origin", "Destination")) |>
  full_join(od_ft_trips_taxi, by=c("Origin", "Destination")) |>
  full_join(od_ft_trips_rideshare, by=c("Origin", "Destination")) |>
  full_join(od_ft_trips_schoolbus, by=c("Origin", "Destination")) |>
  full_join(od_ft_trips_motorcycle, by=c("Origin", "Destination")) |>
  full_join(od_ft_trips_other, by=c("Origin", "Destination")) 
```

Replace all NAs with 0s:
```{r}
od_ft_trips[is.na(od_ft_trips)] <- 0
```

Combine the `od_ft` person data frame with the `od_ft_trips` data frame. Notice that there are more OD pairs in the `od_ft` data frame (`r od_ft$Origin |> length()`) than `od_ft_trips` (`r od_ft_trips$Origin |> length()`) because some OD pairs which have workers and/or are a location for jobs do not have any trips taken that survey day. The TTS 2016 is a travel survey after all. 
```{r}
od_ft_p_t <- full_join(od_ft, od_ft_trips, by=c("Origin", "Destination")) 
```

Check to see if the total number of trips adds up to the total available from TTS 2016 (to ensure we separated and merged all the tables correctly)
```{r}
od_ft_p_t <- od_ft_p_t |> rowwise() |>
  mutate(sum_trips = sum(T_LT,T_LTGO,T_GO,BC,W,C_DPC,C_PPC,C_TA,C_RS,C_SB,M,O, na.rm=T))

#the number of total trips. This value is slightly higher than the value for total trips (not disagg. by mode, which is equal to 3,282,539 - I checked in excel) as available from the TTS 2016. However, I checked the summed up total trip, disaggregated by mode, in excel, and it yields this number of 3,282,611 as well. So there must just be some rounding up that occurred in the TTS 2016 when disaggregating by mode... regardless the numerical difference is insignificant. 
od_ft_p_t$sum_trips |> sum()
```

Replace all NAs with 0s, again:
```{r}
od_ft_p_t[is.na(od_ft_p_t)] <- 0
```

@@

The od table can be used to calculate the count of workers, count of jobs, and count of % transit, bike, and walk slit, by traffic analysis zone:
```{r}
workers_ft <- od_ft_p_t |>
  group_by(Origin) |>
  summarize(workers = sum(Persons),
            .groups = "drop") |>
  transmute(GTA06 = Origin,
            workers)

jobs_ft <- od_ft_p_t |>
  group_by(Destination) |>
  summarize(jobs = sum(Persons),
            .groups = "drop") |>
  transmute(GTA06 = Destination,
            jobs)
```

Active travel modal splits:
```{r}
LTGO_split_ft <- od_ft_p_t |>
  group_by(Origin) |>
  summarize(LTGO_split = sum(T_LT, T_LTGO, T_GO)/sum(sum_trips)) |> 
  transmute(GTA06 = Origin,
            LTGO_split)

W_split_ft <- od_ft_p_t |>
  group_by(Origin) |>
  summarize(W_split = sum(W)/sum(sum_trips)) |>
  transmute(GTA06 = Origin,
            W_split)

BC_split_ft <- od_ft_p_t |>
  group_by(Origin) |>
  summarize(BC_split = sum(BC)/sum(sum_trips)) |>
  transmute(GTA06 = Origin,
            BC_split)
```

Join workers and jobs to simple features object:
```{r}
ggh_taz <- ggh_taz |>
  # Join workers
  left_join(workers_ft,
            by = "GTA06") |>
  # Join jobs
  left_join(jobs_ft,
            by = "GTA06") |>
  left_join(LTGO_split_ft,
            by = "GTA06") |>
  left_join(W_split_ft,
            by = "GTA06") |>
  left_join(BC_split_ft,
            by = "GTA06")
```

Replace NAs with 0s for the new columns:
```{r}
ggh_taz <- ggh_taz |>
  mutate_at(vars(workers, jobs, LTGO_split, W_split, BC_split), ~replace_na(., 0))
```

## Visualize

@@

Plot transit split (0 to 1, 1 being 100% of trips made using transit) distribution in the ggh. Aggregating by CMA could be interesting:
```{r}
ggplot() +
  geom_sf(data = ggh_taz,
          aes(fill = LTGO_split),
          color = NA) +
  scale_fill_distiller(direction = 1)
```
Let's zoom into Toronto CMA, and plot the transit split
```{r}
#rotating the WGS84 crs so it displays Toronto aesthetically on a page; 17degree shift using +gamma
# lat_0 and lonc are the approximate centroid of Toronto
crs_string <- "+proj=omerc +lat_0=43.6532 +lonc=-79.3832 +alpha=0 +k_0=.7 +datum=WGS84 +units=m +no_defs +gamma=17"

#the bounding box, so we focus on Toronto
toronto_bb <- st_bbox(ggh_cma |> filter(CMANAME %in% c("Toronto")))

tm_shape(ggh_taz, bbox = toronto_bb, projection=crs_string) +
  tm_polygons(col = "LTGO_split", border.alpha = 0.1, style = "jenks") +
  tm_shape(ggh_cma |> filter(CMANAME %in% c("Toronto")), bbox = toronto_bb, projection=crs_string) +
  tm_borders()
```
The CMA/CAs overlapping all the TAZ:
```{r}
tm_shape(ggh_taz, projection=crs_string) +
  tm_polygons(col = "CMAUID", border.alpha = 0.1, style = "jenks") +
  tm_shape(ggh_cma, projection=crs_string) +
  tm_borders()
```

The planning boundaries overlapping all the TAZ:
```{r}
tm_shape(ggh_taz, projection=crs_string) +
  tm_polygons(col = "REGION", border.alpha = 0.1, style = "jenks") +
  tm_shape(ggh_pd, projection=crs_string) +
  tm_borders()
```

@@

Attach the centroid point geometry to the origin object (workers) and the destination object (jobs)
```{r}
#workers (origins)
workers_ft_cent <- workers_ft |> merge(ggh_taz_cent |> dplyr::select(GTA06), by="GTA06") |> st_sf()

work_origins <- cbind(workers_ft_cent, st_coordinates(st_transform(workers_ft_cent, crs = 4326))) |>
  dplyr::rename(lon = "X", lat = "Y", id = "GTA06") |> as.data.frame()  |>
  dplyr::select(id, lon, lat) |> as.data.frame()

@@

```{r}
#jobs (destinations)
jobs_ft_cent <- jobs_ft |> merge(ggh_taz_cent |> dplyr::select(GTA06), by="GTA06") |> st_sf

job_destinations <- cbind(jobs_ft_cent, st_coordinates(st_transform(jobs_ft_cent, crs = 4326))) |>
  dplyr::rename(lon = "X", lat = "Y", id = "GTA06") |> as.data.frame()  |>
  dplyr::select(id, lon, lat)

### 2023-11-08 data-raw/01-TTS16-Calculating-Travel-Time-OD.Rmd
Working on development branch to ensure reproducibility of data processing and adding new big od table

Libraries:
```{r setup, include=FALSE}
library(disk.frame)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(httr)
library(knitr)
#install.packages("kableExtra", dependencies = TRUE) #installing kableExtra and dependencies as I was receiving errors
library(kableExtra)
# install.packages("devtools")
# install.packages("Rtools")
# devtools::install_github("https://github.com/hrbrmstr/lodes.git", dependencies = TRUE) # if you need to download
library(lodes)
library(progress)
library(purrr)
library(r5r)
library(sf)
library(stplanr)
library(tidycensus)
library(tidyr)
library(tmap)
library(zoo) # for rollapplyr

# setup options
setup_disk.frame() # setup disk.frame parallel computing
options(future.globals.maxSize = Inf) # allow disk.frame threads to talk to eachother without limits
options(java.parameters = "-Xmx6G") # assign memory for r5r to work with
tmap_mode("plot")
```

>>>>

Libraries:
```{r setup, include=FALSE}
library(disk.frame) # not installed on this machine
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(gridExtra) # Miscellaneous Functions for "Grid" Graphics
library(httr) # Tools for Working with URLs and HTTP
library(knitr) # A General-Purpose Package for Dynamic Report Generation in R
#install.packages("kableExtra", dependencies = TRUE) #installing kableExtra and dependencies as I was receiving errors
library(kableExtra) # Construct Complex Table with 'kable' and Pipe Syntax
# install.packages("devtools")
# install.packages("Rtools")
# devtools::install_github("https://github.com/hrbrmstr/lodes.git", dependencies = TRUE) # if you need to download
#library(lodes) # not installed on this machine
library(progress) # Terminal Progress Bars
library(purrr) # Functional Programming Tools
library(sf) # Simple Features for R
library(stplanr) # not installed on this machine
library(tidycensus) # not installed on this machine
library(tidyr) # Tidy Messy Data
library(tmap) # Thematic Maps
library(zoo) # for rollapplyr

# setup options
setup_disk.frame() # setup disk.frame parallel computing
options(future.globals.maxSize = Inf) # allow disk.frame threads to talk to eachother without limits
tmap_mode("plot")
```

```{r}
options(java.parameters = "-Xmx30G") # assign memory for r5r to work with
library(r5r) # Rapid Realistic Routing with 'R5' 
```
